Trait pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues

### 2022-08-28 ms/ms01_titlepage.tex
ENH added supp table to snakemake

% Verify numbers
First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
%
I have categorized the GWAS traits into 96 traits and categories to define pleiotropic as belonging at least to two of these categories.
%
Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
%
These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
%
These variants with the GWAS traits, egenes and etissues can be explored in a public website.

@@

\begin{keywords}
keyword1 | keyword2 | keyword3
\end{keywords}

\begin{corrauthor}
aitor.gonzalez\at univ-amu.fr
\end{corrauthor}

>>>>

% Verify numbers
First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
%
I have categorized the GWAS traits into 96 traits and categories to define pleiotropic as belonging at least to two of these categories.
%
Pleiotropic variants are associated to more eQTL genes (eGenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
%
These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
%
These variants with the GWAS traits, egenes and etissues can be explored in a public website.

@@

\begin{keywords2}
eGenes; eQTLs
\end{keywords2}

### 2022-08-28 ms/ms02_intro.tex
ENH added supp table to snakemake

Despite these previous disadvantages eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (egenes) and tissues (etissues) (REF).
%
In addition colocalization anlaysis can pinpoint more precisely the causal variants that either data set independently (REF).

The large number of GWAS has allowed to discover that many genetic variants are involved in different traits.
%
TODO explain Watanabe.
%
Because GWAS variants affect mostly non-coding regions, it makes sense to investigate trait pleiotropy in terms of non-coding gene regulatory functions.

For these reasons, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
%
I have used these colocalization results to investigate the gene regulatory properties of pleiotropic variants.
%
This global colocalization analysis is also useful to help interpret GWAS variants.

>>>>

Nevertheless eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (eGenes) and tissues (etissues) (REF).
%
In addition colocalization anlaysis can pinpoint more precisely the causal variants that either data set independently (REF).

The large number of GWAS has allowed to discover that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
%
In this work, the analysed loci covered 60\% of the genome and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
%
They also found that pleiotropic genes and variants where less tissue specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
%
The question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature.Geneticeffectsgene}.

In this work, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
%
I have used these colocalization results to investigate molecular differences between pleiotropic variants in terms of gene regulatory properties.
%
We provide a model to mechanistically understand our results.
%
This global colocalization analysis is also useful to help interpret GWAS variants.
